Simulation of Nano Particles in a Laser Trap

### 2017-03-01 doc/draft4.tex
updates

Miller et al. investigate the dependence of the emerging temperature $T_\text{em}$ on the intensity of the laser for different sized nano particles
and different pressures.\\
With these four temperatures introduced, some questions arise about the relationship between these temperatures. For this theses, I want to focus on
the question: how does the laser intensity influence the motion of the nano particle in the laser trap and its center of mass velocity?\\

@@

distribution. However, the positions on the FCC lattice are not very natural and so the atoms equilibrate to a different temperature than desired.

@@

irregularities (jumps) before the system equilibrates to a constant value. This equilibration process divides the simulation in two phases: the equilibration

@@

        \caption{Temperature and energy in the equilibration phase of the velocity Verlet algorithm. The velocity rescaling (see eq. 

@@

    jumps are happening. The desired temperature in this case is $T^* = 0.2$, which is reached after about 7000 steps.}

@@

\subsection{eHEX}
The next checkpoint is the eHEX algorithm. Since it contains several steps and variables that have to be calculated, this algorithm is very prone
to errors. Perhaps the most crucial of those variables in the algorithm is the amount of heat injected into the system, $\Delta Q$. 
The effect of the value of this variable on the system has therefore to be checked to make sure the system is not overheating.\\
The investigation of the effects on the system can be done in two ways: comparing the temperature and the energy of the system during the 
application of the algorithm (similar to the way it's been done with the velocity Verlet) and comparing the effect of different values of $\Delta Q$.\\
%Like mentioned above, there is an equilibration phase, which takes place before the recording of the values for this comparison. The graph starts 
%with a velocity Verlet phase after which the eHEX algorithm starts, to emphasize the effect of the change in temperature and energy.\\
Since the eHEX algorithm injects heat into the system, one would expect the energy and the temperature to rise during the application of the
algorithm. In Fig.~\ref{fig:ehex_te} we can see that exactly this is happening. The graph shows a velocity Verlet phase (that follows an equilibration
phase which is not recorded) of constant temperature and energy and an eHEX phase, where the temperature and the enrgy are constantly rising. 
\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.4]{images/ehex_te.pdf}
        \caption{Comparison of temperature (red) and energy (blue) during the application of the eHEX algorithm with $\Delta Q=0.04$ (in reduced units).
                The first 20000 timesteps, the energy and temperature are constant during the application of the velocity Verlet algorithm and start
            to rise, once the eHEX algortithm takes over. The values for the temperature can be seen on the ordinate on the left hand side
            and the values for the energy on the right hand side. }
        \label{fig:ehex_te}
    \end{center}
\end{figure}
The comparison of different values of $\Delta Q$ can be seen in Fig.~\ref{fig:ehex_dq}. As the correlation between the temperature and the energy has
been shown in Fig.~\ref{fig:ehex_te}, the plotting of the temperature data should suffice. 
\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.4]{images/ehex_temp_dq.pdf}
        \caption{Comparison of different values for $\Delta Q$, ranging from 0 to 0.09. As the value $\Delta Q=0$ corresponds to no heat being
        injected into the system, it is not surprising that the temperature remains constant. The graph shows that an increase in $\Delta Q$ leads to
    an increase in the rate of warming up, which means that for bigger $\Delta Q$, the system is heating up more quickly.}
        \label{fig:ehex_dq}
    \end{center}
\end{figure}

@@

\subsection{Measurement}

@@

internal temperature of the nano particle (set with the velocity verlet) $T_\text{intern} = 0.2$ and a system size of $N=32$ atoms in the nano

@@

@@

temperatures. The reason for this may be in the difference of the accomodation factor $\alpha$. In the paper, Millen et al. use an accomodation factor

@@

        \includegraphics[scale=0.3]{images/measurevscalc.pdf}
        \caption{Comparison between the measured (red) and calculated (blue) center of mass temperature according to \cite{MillenJ.2014}.}
        \label{fig:measurevscalc}

@@

\newpage
\section{Conclusion and Further Remarks}
% what does this all tell us? do we need to study more of this? 
In this thesis I have modeled an optical tweezer experiment with methods from the field of computational physics. The experiment was broken down into
different sections, that have been modeled separately. With the use of molecular dynamics I simulated the behaviour of a nano particle in a laser trap
and investigated the effects of the laser power and temperature of the surrounding gas on the center of mass movement of the nano particle.\\
The results show, that the center of mass temperature (and thus its movement) does indeed depend on the energy of the laser, as well as on the
pressure and temperature of the surrounding gas.\\
The development of the center of mass temperature shows a curve that seems to have a convex form, whereas the internal temperature shows a more
concave behaviour. So while the internal temperature initially grows faster, it could be possible for the center of mass temperature to pass the
internal temperature.\\
There are several ways to explore this subject further. The first idea would be to decrease the pressure even more and investigate the case of very low
energies, since this is not easily feasible in an experimental setup. Another idea would be to increase the number of atoms in the nano particle to
see if the surface of the particle plays a role in the development of the temperatures. Furthermore, the interaction between the gas particles and the
nano particle in the trap could be varied to include an attractive part in the potential. Generally, these simulations could be expanded by running
them on faster computers to investigate long term behaviour of the systems and expand the ranges of the different values like pressure or laser
energy.

>>>>

As described above, the impining and emerging gas particles form two independent heat baths with two different temperatures 
with $T_\text{em} > T_\text{imp}$ in general. This leads to two different friction coefficients and noise terms that have to be included in the 
Langevin equation.\\
Millen et al. investigate the dependence of the emerging temperature $T_\text{em}$ on the intensity of the laser for different sized nano particles
and different pressures. They show that the emerging gas temperature increases with increasing laser power and depends on the size of the nano
particle, as the emerging gas temperature increases slower for bigger particles.\\
Where the experiment conducted by Gieseler et al. uses only the temperature of the heat bath for their model, Millen et al. establish four different
temperatures that are present in the experiment: the temperature of the impinging and emerging gas particles, the
surface temperature and the center of mass temperature of the nano particle in the trap. Since the center of mass motion of the nano particles
is the main objective in the experiment of Gieseler et al., the dependence of the center of mass temperature on 
laser intensity and the temperature of the impinging gas should be investigated, which will be done in this thesis. The questions I would like to
answer are: How does the laser intensity influence the motion of the nano particle in the laser trap and its center of mass velocity? What impact
does the impinging gas temperature have on the temperature of the center of mass?\\
%With these four temperatures introduced, some questions arise about the relationship between these temperatures. For this theses, I want to focus on
%the question: how does the laser intensity influence the motion of the nano particle in the laser trap and its center of mass velocity?\\

@@

distribution. As it is very unlikely for the atoms to be sitting exactly on a FCC lattice, the first few time steps are needed for the equilibriation
of the system. Due to this process the temperature of the system can change slightly and vary from the desired temperature. 
%However, the positions on the FCC lattice are not very natural and so the atoms equilibrate to a different temperature than desired.

@@

jumps before the system equilibrates to a constant value. This equilibration process divides the simulation in two phases: the equilibration

@@

        \caption{Temperature and energy in the equilibration phase of the nano particle without the gas with a time step of
            $\Delta t = 0.02$. The velocity rescaling (see eq. 

@@

    jumps are happening. The desired temperature in this case is $T^* = 0.2$, which is reached after about 7000 steps. The fluctuation of the energy
    is due to the introduction of the cutoff radius $r_C$ which is set at $r_C = 2.5$.}

@@

\subsection{eHEX}
The next checkpoint is the eHEX algorithm. Since it contains several steps and variables that have to be calculated, this algorithm is very prone
to errors. Perhaps the most crucial of those variables in the algorithm is the amount of heat injected into the system, $\Delta Q$. 
The effect of the value of this variable on the system has therefore to be checked to make sure the system is not overheating.\\
The investigation of the effects on the system can be done in two ways: examining the development of the temperature and the energy of the system 
during the 
application of the algorithm (similar to the way it's been done with the velocity Verlet) and comparing the effect of different values of $\Delta Q$.\\
%Like mentioned above, there is an equilibration phase, which takes place before the recording of the values for this comparison. The graph starts 
%with a velocity Verlet phase after which the eHEX algorithm starts, to emphasize the effect of the change in temperature and energy.\\
Since the eHEX algorithm injects heat into the system, one would expect the energy and the temperature to rise during the application of the
algorithm. In Fig.~\ref{fig:ehex_te} we can see that exactly this is happening. The graph shows a velocity Verlet phase (that follows an equilibration
phase which is not recorded) of constant temperature and energy and an eHEX phase, where the temperature and the enrgy are constantly rising. 
\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.4]{images/ehex_te.pdf}
        \caption{Comparison of temperature (red) and energy (blue) during the application of the eHEX algorithm with $\Delta Q=0.04$ (in reduced units).
                The first 20000 timesteps, the energy and temperature are constant during the application of the velocity Verlet algorithm and start
            to rise, once the eHEX algortithm takes over. The values for the temperature can be seen on the ordinate on the left hand side
            and the values for the energy on the right hand side. }
        \label{fig:ehex_te}
    \end{center}
\end{figure}
The comparison of different values of $\Delta Q$ can be seen in Fig.~\ref{fig:ehex_dq}. As the correlation between the temperature and the energy has
been shown in Fig.~\ref{fig:ehex_te}, the plotting of the temperature data should suffice. 
\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.4]{images/ehex_temp_dq.pdf}
        \caption{Comparison of different values for $\Delta Q$, ranging from 0 to 0.09. Up to time $t = 200$ only the velocity Verlet algorithm is
            applied which is followed by the application of the eHEX algorithm at different values for $\Delta Q$.
            As the value $\Delta Q=0$ corresponds to no heat being
        injected into the system, it is not surprising that the temperature remains constant. The graph shows that an increase in $\Delta Q$ leads to
    an increase in the rate of warming up, which means that for bigger $\Delta Q$, the system is heating up more quickly.}
        \label{fig:ehex_dq}
    \end{center}
\end{figure}

@@

\subsection{Simulation of the Experiment}

@@

internal temperature of the nano particle $T_\text{intern} = 0.2$ and a system size of $N=32$ atoms in the nano

@@

    \label{eq:compaper}

@@

temperatures. The reason for this may be in the difference of the accomodation factor $\alpha$, which is given by~\cite{Goodman1980}
    \alpha = \frac{T_\text{em} - T_\text{imp}}{T_\text{sur} - T_\text{imp}}
In the paper, Millen et al. use an accomodation factor

@@

The thermal accomodation factor quantifies the behaviour of the interaction between the gas particles and the nano particle. Although Eq.
\eqref{eq:compaper} does not contain an explicit dependence on the thermal accomodation factor, the difference between the value in the experiment and
in the simulation suggests a difference in the energy transfer during interaction between the gas and the nano particle.

\begin{figure}
    \begin{center}
        \includegraphics[scale=0.3]{images/measurevscalc.pdf}
        \caption{Comparison between the measured (red) and calculated (blue) center of mass temperature according to \cite{MillenJ.2014}.}
        \label{fig:measurevscalc}
    \end{center}
\end{figure}

@@

\newpage
\section{Conclusion and Further Remarks}
% what does this all tell us? do we need to study more of this? 
In this thesis I have modeled an optical tweezer experiment with methods from the field of computational physics. The experiment was broken down into
different sections, that have been modeled separately. With the use of molecular dynamics I simulated the behaviour of a nano particle in a laser trap
and investigated the effects of the laser power and temperature of the surrounding gas on the center of mass movement of the nano particle.\\
The results show, that the center of mass temperature (and thus its movement) does indeed depend on the energy of the laser, as well as on the
pressure and temperature of the surrounding gas.\\
The development of the center of mass temperature shows a curve that seems to have a convex form, whereas the internal temperature shows a more
concave behaviour. So while the internal temperature initially grows faster, it could be possible for the center of mass temperature to pass the
internal temperature.\\
There are several ways to explore this subject further. The first idea would be to decrease the pressure even more and investigate the case of very low
energies, since this is not easily feasible in an experimental setup. Another idea would be to increase the number of atoms in the nano particle to
see if the surface of the particle plays a role in the development of the temperatures. Furthermore, the interaction between the gas particles and the
nano particle in the trap could be varied to include an attractive part in the potential. Generally, these simulations could be expanded by running
them on faster computers to investigate larger systems and expand the ranges of the involved values like pressure, laser
energy, and internal temperature.
